---
title: "Motortrend Study"
author: "Harland Hendricks"
date: "August 7, 2018"
output: pdf_document
---
##Executive Summary

*Motor Trend* magazine is interested in exploring the relationship between a set
of variables from a data set of a collection of cars and miles per gallon (MPG).

Specifically, the magazine would like to focus on the following two questions:

* "Is an automatic or manual transmission better for MPG"

* "Quantify the MPG difference between automatic and manual transmissions"

When testing the predictors **wt**, **cyl**, and **am** using the nested model method, 
I found that **am** is not significant and should not be used to model affects on MPG.  A simple plot of MPG per
type of transmission is all that is required to determine that a manual transmission 
can result in higher gas mileage for the data set used.  Additionally, simply subtracting
the mean of automatic transmission MPG from manual transmission MPG quantifies the 
difference in MPG.

An appropriate model for MPG for the given data is one that includes the predictors 
**wt** and **cyl**.  These predictors appear to influence MPG more than **am**.
This model suggests that for every 1000 lbs increase in the automobile weight, MPG 
decreases by 3.2 MPG.  Additionally, MPG will decrease by 4.3 when increasing cylinders 
from 4 to 6, and will decrease by 6.1 when increasing cylynders from 4 to 8.

##mtcars Dataset

The instructions for this assignment are located [here.](https://www.coursera.org/learn/regression-models/peer/nxntd/regression-models-course-project)

Built with R version `r getRversion()` with the following system:

```{r, echo = FALSE}
Sys.info()
```

Load the required libraries

```{r}
library(dplyr)
library(ggplot2)
```

Explore mtcars dataset

```{r}
head(mtcars, 3)
```

```{r}
str(mtcars)
```

Create *data* object with **mpg**, **cyl**, **wt**, and **am** variables.  Reclassify **cyl** and **am** 
from numeric to factor and add column **trans** to indicate transmission type.

```{r}
data <- mtcars[, c(1, 2, 6, 9)]
data$cyl <- as.factor(data$cyl)
data$am <- as.factor(data$am)
data <- mutate(data, trans = ifelse(am == 1, "manual", "auto"))
```

Plot **trans** vs **mpg**

```{r, fig.height = 4, fig.width = 4, echo = FALSE}
g <- ggplot(data, aes(wt, mpg, color = cyl))
g + geom_point() +
      facet_grid(. ~ trans) +
      ggtitle("Miles per Gallon (MPG) per Weight, Number of Cylinders, and Transmission Type") +
      labs(x = "Weight (1000 lbs)", y = "MPG") +
      theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "in"))
```

The initial plot of MPG given a weight, number of cylinders, and specific transmission type 
indicates that as weight increases - mpg decreases.  Additionally, as the number of cylinders 
increases - mpg decreases.  Finally, a manual transmission provides greater miles per gallon.

##Model Selection and Fit

I chose to use the predictors **cyl**, **wt**, and **am** due to the belief that
these variables would have the most affect on **mpg**.  *Motor Trend* is also specifically
interested in **am**.  I will build three models and then nest the models to determine which 
model contains necessary predictors.

###Model 1

```{r}
m1 <- lm(mpg ~ wt, data = data)
summary(m1)
```

We can then plot the residuals for diagnostics:

```{r, fig.height=4, fig.width=4}
par(mfrow = c(2, 2))
plot(m1)
```

The top left and top right residual plots indicate that there may be a problem modeling
with just **wt**

##Model 2

```{r, echo = FALSE}
m2 <- lm(mpg ~ wt + cyl, data = data)
m2$call
m2$coefficients
```

We can then plot the residuals for diagnostics:

```{r, fig.height=4, fig.width=4, echo = FALSE}
par(mfrow = c(2, 2))
plot(m2)
```

This model looks better, but the top right residual plot still indicates that the residuals
may not be normally distributed and there may be a problem modeling with **wt** and **cyl**.

##Model 3

```{r, echo = FALSE}
m3 <- lm(mpg ~ wt + cyl + am, data = data)
summary(m3)
```

We can then plot the residuals for diagnostics:

```{r, fig.height=4, fig.width=4, echo = FALSE}
par(mfrow = c(2, 2))
plot(m3)
```

The residual plots for model 3 don't change much from 2, but the p values for **am** are 
not significant.  There may be a problem modeling with **wt**, **cyl**, and **am**.

##Nested Model Testing

```{r}
anova(m1, m2, m3)
```


Using the nested model test, I determine that the model with predictors **wt** and
**cyl** is significant with acceptable residual plots.

The predictor **am** is not significant and should not be used to model **mpg** in
*Motor Trend's* study.  The linear model with **am** as a predictor just predicts the 
mean of **mpg** when grouped by manual or automatic transmission and is not a very useful
model

The mean MPG of each **am** show that a manual transmission is better for **mpg** and
that the difference between manual and automatic translates to a difference in MPG:

```{r, fig.height=3, fig.width=3, echo = FALSE}
g2 <- ggplot(data, aes(trans, mpg))
g2 + geom_boxplot() +
      ggtitle("Miles per Gallon (MPG) per Type of Transmission") +
      labs(x = "Transmission Type", y = "MPG") +
      theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "in"))
```
